Plot density of imputations

### 2020-02-20 3.Thesis/plot_imputation_densities.Rmd
re-wrote code to develop plots for all variables, not only bmi

---
title: "Plot density of imputations"
output: html_notebook
---

```{r echo=TRUE, message=FALSE, warning=FALSE}
# set-up environment
library(mice) # Data imputation
library(dplyr) # Data manipulation
# library(magrittr)
# library(broom)
library(ggplot2)
# library(purrr)
set.seed(123)
dat <- nhanes
mids <- list()

# impute missingness with one iteration
M <- 20
mids <- mice(dat, maxit = 1, m = M, print = FALSE)

####
# this should be repeated T times
####
# add an iteration
mids <- mice.mids(mids, maxit = 1, printFlag = FALSE)
# # traceplots
# plot(mids)

# plot density of imputed values
imputed_values <- mids[["imp"]]$bmi %>% mutate(case = row_number()) %>%  tidyr::gather("imp", "value", 1:M) %>% .[order(.$case, .$imp), ] #%>% group_by(case) %>% mutate(case_mean = mean(value)) %>% ungroup()
case_mean <- imputed_values %>% group_by(case) %>% summarise(mu = mean(value)) 
# # for one case
# imputed_values %>% filter(case==3) %>%  ggplot() +
#   geom_histogram(aes(x = value)) + 
#   geom_vline(xintercept = mean(imputed_values$value))

# # store names of cases
# missing_cases <- 1:max(imputed_values$case)#unique(imputed_values$case)
# # plot histograms
# missing_cases %>%
#   purrr::map(function(z) {imputed_values %>% filter(case==z) %>% ggplot() +
#   geom_histogram(aes(x = value)) 
#   })

# plot hist per case
imputed_values %>% ggplot(aes(x = value)) + 
  geom_histogram() +
  geom_vline(data = case_mean, mapping = aes(xintercept = mu)) +
  facet_wrap(~case)

@@

>>>>

---
title: "Plot density of imputations"
output:
  html_document:
    df_print: paged
---

```{r echo=TRUE, message=FALSE, warning=FALSE}
# set-up environment
library(mice) # Data imputation
library(dplyr) # Data manipulation
# library(magrittr)
# library(broom)
library(ggplot2)
library(purrr)
set.seed(123)
dat <- nhanes

# parameters
M <- 20
T_max <- 10

@@

```{r}
# for BMI only
# objects
plot <- list()
distr <- list()

# this should be repeated T times
for (it in 1:T_max) {
  if (it == 1) {
    # impute missingness with one iteration
    mids <- mice(dat,
                 maxit = 1,
                 m = M,
                 print = FALSE)
  } else {
    # add an iteration
    mids <- mice.mids(mids, maxit = 1, printFlag = FALSE)
  }

  # plot density of imputed values
  imputed_values <-
    mids[["imp"]]$bmi %>% mutate(case = row_number()) %>%  tidyr::gather("imp", "value", 1:M) %>% .[order(.$case, .$imp),] #%>% group_by(case) %>% mutate(case_mean = mean(value)) %>% ungroup()
  distr[[it]] <-
    case_mean <-
    imputed_values %>% group_by(case) %>% summarise(mu = mean(value), sigma = sd(value))

  # plot hist per case
  plot[[it]] <- imputed_values %>% ggplot(aes(x = value)) +
    geom_histogram() +
    geom_vline(data = case_mean, mapping = aes(xintercept = mu)) +
    facet_wrap( ~ case)
  # would be nice to have these overlapping per case!

  # # plot density per case
  # imputed_values %>% ggplot(aes(x = value)) +
  #   geom_histogram(aes(y = ..density..)) +
  #   geom_density() +
  #   geom_vline(data = case_mean, mapping = aes(xintercept = mu)) +
  #   facet_wrap(~case)

  ############################ make it work for all other vars
  # vars <- map_dfr(dat, ~is.na(.x) %>% sum() > 0)
  # imputed_values_list <- mids[["imp"]] %>% map(~{filter(.x %in% vars)})
  #
  # #filter(length(row.names(.x)))}) %>% map(~{mutate(case = row_number()) %>%  tidyr::gather("imp", "value", 1:M) %>% .[order(.$case, .$imp), ]})
  ############################
}

# # this is just the between chain variance, right?
# average_sd <- distr %>% purrr::map(~{mean(.$sigma)})
# plot.ts(average_sd) #is this == plot(mids)?

# # now split by case abd plot time series
# purrr::map(1:9, function(x) purrr::map(distr, ~{.$sigma[.$case==x]})) %>% purrr::map(~plot.ts(.x))

# or with ggplot to put them in 1 plot
sd_per_case_per_it <-
  purrr::map(1:9, function(x)
    purrr::map_dbl(distr, ~ {
      .$sigma[.$case == x]
    })) %>% as.data.frame() %>% mutate(nr_it = row_number())

colnames(sd_per_case_per_it) <- c(1:9, "nr_it")

#sd_per_case_per_it %>%  tidyr::gather("case", "sd", 1:9) %>% ggplot() + geom_line(aes(x = nr_it, y = sd)) + facet_wrap(~case)
sd_per_case_per_it %>%  tidyr::gather("case", "sd", 1:9) %>% ggplot(aes(x = nr_it, y = sd, col = case)) + geom_line() + geom_smooth(se = F)

Now for any variable
```{r}
# objects
imputed_values <- plot <- distr <- list()

# which vars?
nmis <- dat %>% map_dbl(~{is.na(.x) %>% sum()}) #%>% which(.>0) %>% names()
#nmis <- apply(is.na(dat), 2, sum)
vars <- names(nmis[nmis>0])

# this should be repeated T times
for (it in 1:T_max) {
  if (it == 1) {
    # impute missingness with one iteration
    mids <- mice(dat,
                 maxit = 1,
                 m = M,
                 print = FALSE)
  } else {
    # add an iteration
    mids <- mice.mids(mids, maxit = 1, printFlag = FALSE)
  }

  # plot density of imputed values
imputed_values[[it]] <-  mids[["imp"]][vars]
}

# #test3 <- map(1:T_max, function(x) map(imputed_values, function(y) y[vars]))
# test4 <- map(imputed_values, function(x) x[vars])

imputed_values[[1]]$bmi %>% mutate(case = row_number())
#map(imputed_values, function(x) map(vars, function(y) {print(x)}))#mutate(case = row_number())})) #%>%  tidyr::gather("imp", "value", 1:M) %>% .[order(.$case, .$imp),]}) 
test5 <- map(imputed_values, function(x) map(x, function(y) {as.data.frame(y) %>% mutate(case = row_number()) %>%  tidyr::gather("imp", "value", 1:M)}))  #%>% .[order(.$case, .$imp),]}) })) 
# case_mean <- imputed_values %>% group_by(case) %>% summarise(mu = mean(value), sigma = sd(value))

# # plot hist per case
# test4[[1]] %>% map(~{ggplot(aes(x = value)) +
#     geom_histogram() +
#     geom_vline(data = case_mean, mapping = aes(xintercept = mu)) +
#     facet_wrap( ~ case)})
